---
title: "Project2_Rmd"
author: "Megan Ball"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages
library(dplyr)
library(readr)
library(gt)
library(summarytools)
library(GGally)
```

# Load the Data 
```{r}
df <- read_csv(here::here("data", "CaseStudy2-data.csv"))

# Check for missing values
tibble(variable = names(colSums(is.na(df))),
       missing = colSums(is.na(df))) %>% 
  gt() %>% 
  tab_header(title = "Missing Values in Data") 

#remove ID
df <- df %>% select(-c(ID))

```



```{r}

#summarize data
summary(df)
print(dfSummary(df, graph.magnif = 0.75), method = 'browser')
str(df)

```

Comments on the data:
- Investigate employee count- it has only one distinct value
- Monthly income is skewed, as is expected with most income data
- Over18 is all Y, so may remove the column as it is not useful data
- Standard Hours is all 80, so may also remove this column
- Investigate monthly income values, $19,999 seems pretty high for the max
- There are no missing values in the data

```{r}
#update all characters to factors
df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)], 
                                       as.factor)

str(df)
```

# Exploratory Data Analysis

```{r include=FALSE}
df$EmployeeCount
```


```{r}
#remove unnecessary values, employee count, over18, and standard hours
df <- df %>% select(-c(EmployeeCount, Over18, StandardHours))
```

We are now down to 32 variables instead of original 35 by removing ones that contribute no value.

```{r}
#Explore relationships with scatterplot matrix

#continuous variables only
df.numeric <- df[ , sapply(df, is.numeric)]
#add back in Attrition
df.numeric <- cbind(df.numeric, df$Attrition)
ggpairs(df.numeric, aes(colour=df$Attrition))

```

```{r}
#categorial variables only
df.cat <- df[, sapply(df, is.factor)]

ggpairs(df.cat, aes(colour=Attrition))
```
```{r}
#Need to look at one by one


```