---
title: "Project2_Rmd"
author: "Megan Ball"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages
library(dplyr)
library(readr)
library(gt)
library(summarytools)
```

# Load the Data 
```{r}
df <- read_csv(here::here("data", "CaseStudy2-data.csv"))

# Check for missing values
tibble(variable = names(colSums(is.na(df))),
       missing = colSums(is.na(df))) %>% 
  gt() %>% 
  tab_header(title = "Missing Values in Data") 

#remove ID
df <- df %>% select(-c(ID))

```



```{r}

#summarize data
summary(df)
print(dfSummary(df, graph.magnif = 0.75), method = 'browser')
str(df)

```

Comments on the data:
-Investigate employee count- it has only one distinct value
-Monthly income is skewed, as is expected with most income data
-Over18 is all Y, so may remove the column as it is not useful data
-Standard Hours is all 80, so may also remove this column
-Investigate monthly income values, $19,999 seems pretty high for the max

# Exploratory Data Analysis

```{r include=FALSE}
df$EmployeeCount

#remove unnecessary values, employee count, over18, and standard hours
df <- df %>% select(-c(EmployeeCount, Over18, StandardHours))

```
We are now down to 32 variables instead of original 35 by removing ones that contribute no value.

```{r}

```

